---
title: "Anoxia vs Heart study with ash"
author: "Kushal K Dey"
date: "6/9/2017"
output: html_document
---

In this script, we look at how anoxia and heart or cardiac diseases can be compared in the context of using lfsr for the adaptive shrinkage on the tissues for each gene. Adaptive shrinkage is performed using a normal distribution centered about 0. 

```{r}
ash_cor_mat <- get(load("../output/ash_cor_only_voom_pearson_halfuniform_gtex_tissues.rda"))
```

```{r}
common_samples <- get(load("../output/common_samples.rda"))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

#library(data.table)
#data <- data.frame(fread("../data/GTEX_V6/cis_gene_expression.txt"))
#matdata <- t(data[,-c(1,2)])

gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))


U <- unique(tissue_labels)
```

```{r echo=FALSE, eval=TRUE}
order_index <- c();
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```


```{r}
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
betahat_sebetahat <- get(load("../output/anoxia_v_heart_cor_ash.rda"))
sample_attributes <- get(load("../output/sample_attributes_filtered.rda"))
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
```

```{r}
betahat_mat <- betahat_sebetahat$post.mean
sebetahat_mat <- betahat_sebetahat$post.sd
lfsr_mat <- betahat_sebetahat$post.lfsr
```

##  lfsr threshold 0.05

We look at which genes have the most number of lfsr values below a chosen threshold.

```{r}
num_lfsr_below_thresh <- apply(lfsr_mat, 2, function(x) return(length(which(x < 0.05))))

imp_genes <- order(num_lfsr_below_thresh, decreasing = TRUE)[1:100]
num_lfsr_below_thresh[imp_genes]

names <- colnames(betahat_mat)[imp_genes]
```

```{r}
write.table(names, file = "../utilities/anoxia_heart/anoxia_heart_genes_ash_0_05.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
out <- mygene::queryMany(names,  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
cbind.data.frame(out$name, out$summary, out$query)
```

```{r}
pathway2 <- read.delim("../utilities/anoxia_heart/anoxia_heart_genes_0_05_pathways.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

We investigate into a little more detail on anoxia and heart diseases.

```{r}
cause_of_death <- sample_attributes$DTHCOD
anoxia_indices <- grep("anoxia", tolower(cause_of_death))
heart_indices_1 <- grep("cardi", tolower(cause_of_death))
heart_indices_2 <- grep("heart", tolower(cause_of_death))
heart_indices <- union(heart_indices_1, heart_indices_2)
```

Time of death for the anoxia ans heart problems
```{r}
sample_attributes$DTHTIME[anoxia_indices]
```

```{r}
sample_attributes$DTHTIME[heart_indices]
```

```{r}
times <- strsplit(as.character(sample_attributes$DTHTIME), "[:]")
time_stamp <- unlist(lapply(times, function(x){
  y <- as.numeric(x[1])
  z <- as.numeric(x[2])
  w <- y*60+z
  return(w)
}))/(24*60)

times_1 <- time_stamp
```

```{r}
plot(density(times_1[anoxia_indices], na.rm = TRUE), main = "Anoxia death times distribution")
plot(density(times_1[heart_indices], na.rm=TRUE), main = "Heart death times distribution")
```

```{r}
sample_attributes$DTHMNNR[anoxia_indices]
```

We now look at the top 2 genes we found from above.

## ENSG00000166189 or HPS6

Check the read expression profile in GTEx here 

[ENSG00000166189](https://gtexportal.org/home/gene/ENSG00000166189)

```{r}
betahat_sebetahat_2 <- get(load("../output/anoxia_v_heart_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
```

```{r}
indices <- match("ENSG00000166189", colnames(betahat_mat))
```


```{r}
plot(tmat[, indices], col="red")
abline(0,0)
```

Which are the tissues with lowest lfsr 

```{r}
sort(lfsr_mat[,indices], decreasing = FALSE)[1:5]
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$DTHCOD
  total_indices <- c(anoxia_indices, heart_indices)
  temp11 <- temp1[total_indices]
  temp21 <- temp2[total_indices]
  fac_manner_death <- c(rep("Anoxia", length(anoxia_indices)),
                      rep("Heart", length(heart_indices)))
  plot(temp11, col=factor(fac_manner_death), ylab = dimnames(person_tissue_genes)[[2]][j],
       ylim = c(2,7))
}
```

```{r}
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(ash_cor_mat[order_index,order_index, indices]),
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
```

## ENSG00000198730 or CTR9

Check the read expression profile in GTEx here 

[ENSG00000198730](https://gtexportal.org/home/gene/ENSG00000198730)

```{r}
betahat_sebetahat_2 <- get(load("../output/anoxia_v_heart_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
```

```{r}
indices <- match("ENSG00000198730", colnames(betahat_mat))
```


```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$DTHCOD
  total_indices <- c(anoxia_indices, heart_indices)
  temp11 <- temp1[total_indices]
  temp21 <- temp2[total_indices]
  fac_manner_death <- c(rep("Anoxia", length(anoxia_indices)),
                      rep("Heart", length(heart_indices)))
  plot(temp11, col=factor(fac_manner_death), ylab = dimnames(person_tissue_genes)[[2]][j],
       ylim = c(2,7))
}
```

```{r}
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(ash_cor_mat[order_index,order_index, indices]),
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
```

