---
title: "sinusoidal patterns Liver"
author: "Kushal K Dey"
date: "7/1/2017"
output: html_document
---

In this script, we fit sinusoidal graphs to the liver expression data for genes with  liver specific effects (from MASH analysis) and Circadian genes in relation to time of death and expression in the GTEx data.

## Lver specific genes 

The liver genes

```{r}
library(cellcycleR)
liver_genes <- as.character(read.table("../utilities/circadian_mash_3/liver.txt")[,1])
```


```{r}
data <- get(load("../output/cellcyler_data_liver_mash.rda"))

times <- data$times
cycle_data <- data$data
G <- dim(cycle_data)[2]
numcells <- dim(cycle_data)[1];
sigma <- array(0,G);
amp <- array(0,G);
phi <- array(0,G);
cell_times_iter <- times
```

```{r}
lmfit_list <- lapply(1:G, function(g)
{
  temp1 <- scales::rescale(cycle_data[,g], to=c(0,1))
  fit1 <- lm(temp1  ~ sin(0.5*cell_times_iter) + cos(0.5*cell_times_iter) -1);
  #plot(cell_times_iter, fit1$fitted.values)
  #points(cell_times_iter, temp1, col="blue")

  temp2 <- scales::rescale(cycle_data[,g], to = c(-1, 0))
  fit2 <- lm(temp2  ~ sin(0.5*cell_times_iter) + cos(0.5*cell_times_iter) -1);
  #plot(cell_times_iter, fit2$fitted.values)
  #points(cell_times_iter, temp2, col="blue")
  s1 <- summary(fit1)$r.squared
  s2 <- summary(fit2)$r.squared
  if( s1 > s2 ){
    fit <- fit1
    scale <- 0
  }else{
    fit <- fit2
    scale <- 1
  }

  out_sigma <- sd(fit$residuals);
  beta1 <- fit$coefficients[1];
  beta2 <- fit$coefficients[2];
  if(beta1==0 & beta2==0){
    stop(paste0("You have a gene with all 0 counts at gene",g));
  }
  out_amp <- sqrt(beta1^2 + beta2^2);
  out_phi <- atan3(as.numeric(beta2), as.numeric(beta1));
  ll <- list("out_amp"=out_amp, "out_phi"=out_phi, "out_sigma"=out_sigma,
             "out_scale" = scale)
  return(ll)
})

```

```{r}
amp <- as.numeric(unlist(lapply(1:length(lmfit_list), function(n) return(lmfit_list[[n]]$out_amp))));
phi <- as.numeric(unlist(lapply(1:length(lmfit_list), function(n) return(lmfit_list[[n]]$out_phi))));
sigma <- as.numeric(unlist(lapply(1:length(lmfit_list), function(n) return(lmfit_list[[n]]$out_sigma))));
scale <- as.numeric(unlist(lapply(1:length(lmfit_list), function(n) return(lmfit_list[[n]]$out_scale))));
```

```{r}
cycle_data_scaled <- matrix(0, dim(cycle_data)[1], dim(cycle_data)[2])
for(l in 1:dim(cycle_data_scaled)[2]){
  if(scale[l] == 0){
    cycle_data_scaled[,l] <- scales::rescale(cycle_data[,l], to=c(0,1))
  }else{
    cycle_data_scaled[,l] <- scales::rescale(cycle_data[,l], to=c(-1,0))
  }
}
```

```{r}
fitted_val <- matrix(0, dim(cycle_data)[1], dim(cycle_data)[2])
for(g in 1:G){
  fitted_val[,g] <- amp[g]*sin(0.5*cell_times_iter + phi[g])
}
```

```{r}
par(mfrow=c(3,3))
for(g in 1:200){
  plot(times, fitted_val[,g])
  points(times, cycle_data_scaled[,g], col="blue", pch=20)
}
```

```{r}
up_genes <- liver_genes[which(scale == 1)]
down_genes <- liver_genes[which(scale == 0)]
write.table(up_genes, file = "../utilities/circadian_mash_3/up_liver_genes.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(down_genes, file = "../utilities/circadian_mash_3/down_liver_genes.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r fig.height = 10}
pathway2 <- read.delim("../utilities/circadian_mash_3/up_liver_pathway_1.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.5, cexCol = 0.5, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```


```{r fig.height = 10}
pathway2 <- read.delim("../utilities/circadian_mash_3/up_liver_pathway_2.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.5, cexCol = 0.5, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```


```{r fig.height = 10}
pathway2 <- read.delim("../utilities/circadian_mash_3/down_liver_pathway_2.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 1, cexCol = 1, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Circadian genes 

We carry out a similar analysis, but now for the Circadian genes.

```{r}
gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))


pathways <- read.delim("../utilities/CPDB_pathways_genes.tab")
pathway_names <- pathways[,1]
grep("Circadian",pathway_names)
circadian_pathways <- pathways[grep("Circadian", pathway_names),]
circadian_genes <- list()
for (i in 1:dim(circadian_pathways)[1]){
  circadian_genes[[i]] <- strsplit(as.character(circadian_pathways[i,4]), "[,]")[[1]]
}

circ_names <- Reduce(union, circadian_genes)

indices1 <- match(as.character(circ_names), gene_names_1)
indices2 <- indices1[!is.na(indices1)]

```

```{r}
data <- get(load("../output/cellcyler_data_liver_circadian.rda"))

times <- data$times
cycle_data <- data$data
G <- dim(cycle_data)[2]
numcells <- dim(cycle_data)[1];
sigma <- array(0,G);
amp <- array(0,G);
phi <- array(0,G);
cell_times_iter <- times
```

```{r}
lmfit_list <- lapply(1:G, function(g)
{
  temp1 <- scales::rescale(cycle_data[,g], to=c(0,1))
  fit1 <- lm(temp1  ~ sin(0.5*cell_times_iter) + cos(0.5*cell_times_iter) -1);
  #plot(cell_times_iter, fit1$fitted.values)
  #points(cell_times_iter, temp1, col="blue")

  temp2 <- scales::rescale(cycle_data[,g], to = c(-1, 0))
  fit2 <- lm(temp2  ~ sin(0.5*cell_times_iter) + cos(0.5*cell_times_iter) -1);
  #plot(cell_times_iter, fit2$fitted.values)
  #points(cell_times_iter, temp2, col="blue")
  s1 <- summary(fit1)$r.squared
  s2 <- summary(fit2)$r.squared
  if( s1 > s2 ){
    fit <- fit1
    scale <- 0
  }else{
    fit <- fit2
    scale <- 1
  }

  out_sigma <- sd(fit$residuals);
  beta1 <- fit$coefficients[1];
  beta2 <- fit$coefficients[2];
  if(beta1==0 & beta2==0){
    stop(paste0("You have a gene with all 0 counts at gene",g));
  }
  out_amp <- sqrt(beta1^2 + beta2^2);
  out_phi <- atan3(as.numeric(beta2), as.numeric(beta1));
  ll <- list("out_amp"=out_amp, "out_phi"=out_phi, "out_sigma"=out_sigma,
             "out_scale" = scale)
  return(ll)
})

```

```{r}
amp <- as.numeric(unlist(lapply(1:length(lmfit_list), function(n) return(lmfit_list[[n]]$out_amp))));
phi <- as.numeric(unlist(lapply(1:length(lmfit_list), function(n) return(lmfit_list[[n]]$out_phi))));
sigma <- as.numeric(unlist(lapply(1:length(lmfit_list), function(n) return(lmfit_list[[n]]$out_sigma))));
scale <- as.numeric(unlist(lapply(1:length(lmfit_list), function(n) return(lmfit_list[[n]]$out_scale))));
```

```{r}
cycle_data_scaled <- matrix(0, dim(cycle_data)[1], dim(cycle_data)[2])
for(l in 1:dim(cycle_data_scaled)[2]){
  if(scale[l] == 0){
    cycle_data_scaled[,l] <- scales::rescale(cycle_data[,l], to=c(0,1))
  }else{
    cycle_data_scaled[,l] <- scales::rescale(cycle_data[,l], to=c(-1,0))
  }
}
```

```{r}
fitted_val <- matrix(0, dim(cycle_data)[1], dim(cycle_data)[2])
for(g in 1:G){
  fitted_val[,g] <- amp[g]*sin(0.5*cell_times_iter + phi[g])
}
```

```{r}
par(mfrow=c(3,3))
for(g in 1:G){
  plot(times, fitted_val[,g])
  points(times, cycle_data_scaled[,g], col="blue", pch=20)
}
```

```{r}
circ_genes <- colnames(cycle_data)
up_genes <- circ_genes[which(scale == 1)]
down_genes <- circ_genes[which(scale == 0)]
write.table(up_genes, file = "../utilities/circadian_mash_3/up_circ_genes.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(down_genes, file = "../utilities/circadian_mash_3/down_circ_genes.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

We do not find all Circadian genes to have a sinusoidal curve but some of them do have it. But here also, you can see that the peaks of the sine curves are more or less in the same position. This is the problem with this, since in cellcycleR, we draw information from different sets of genes having different phases, which removes the identifiability problem.

