---
title: 'GTEx data: can we compare CorShrink with corpcor and Glasso?'
author: "Kushal K Dey"
date: "5/12/2017"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

Corhrink is an ideal choice for shrinking tissue-tissue correlation matrix
in GTEx data, because these correlations are computed over varying number of
samples for different tissues. `corpcor` is designed to take a data matrix 
as input, whereas the input for glasso is a covariance matrix. So, how do we
use corpcor in this case to compare with corshrink?

So, if one computes a person by tissues by genes matrix, there will be many NAs 
corresponding to cells where a person did not contribute a tissue for all genes.
In order to empower corpcor, I decided to simulate the NA values per person per 
gene from the non-NA values in the tissues the person contributed.

The idea behind doing this is that if a person has given very few tissues, 
we are going to assign a bunch of random samples in the NA places which will bring
down the the correlation between tissues that this person contributes to. As number of persons with NA for a tissue grows, there is a likely chance that the correlation will shrink more and more towards zero.

First we look at how the person by tissue table looks like.

```{r}
person_label=read.table("../data/GTEX_V6/person_identifier_labels_with_numbers.txt");
samples_id <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,1]

samples_person <- sapply(samples_id, function(x) return(paste0(strsplit(as.character(x), "-")[[1]][1:2], collapse ="-")))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

table <- xtabs(~ samples_person + tissue_labels)

unique_persons <- unique(samples_person)
unique_tissues <- unique(tissue_labels)

```

The highest number of tissues contributed by a person is 

```{r}
max(rowSums(table))
unique_persons[which.max(rowSums(table))]
```

The tissue with most number of contributions is 

```{r}
max(colSums(table))
which.max(colSums(table))
```

Next we build a person by tissue by gene matrix.

```{r echo=TRUE, eval=FALSE}

library(data.table)
data <- data.frame(fread("../data/GTEX_V6/cis_gene_expression.txt"))
matdata <- t(data[,-c(1,2)])

mat <- matrix(0, length(unique_persons), length(unique_tissues))
person_tissue_genes <- array(0, c(length(unique_persons), length(unique_tissues), dim(matdata)[2]))

for(u in 1:length(unique_persons)){
    index_samp <- which(samples_person == samples_person[u])
    tissue_lab_samp <- tissue_labels[index_samp]
    person_tissue_genes[u, match(tissue_lab_samp, unique_tissues), ] = matdata[index_samp, ]
    cat("We are at person : ", u, "\n")
}

save(person_tissue_genes, file = "../output/person_tissue_genes.rda")
```

We load the matrix 

```{r}
person_tissue_genes <- get(load("../output/person_tissue_genes.rda"))
```

Next we choose a gene for which we want to compare the tissue correlation 
structure.

```{r}
numg <- 100
```

We fill in the NAs by resampling and then take the ranks of the observations for
robust inference (getting rid of the library size issues)

```{r}
mat <- person_tissue_genes[,,numg]
rank_mat <- matrix(0, dim(mat)[1], dim(mat)[2])
for(u in 1:dim(mat)[1]){
  temp <- mat[u, ]
  w <- which(temp == 0);
  mat[u, w] = sample(temp[temp != 0], length(w), replace=TRUE)
  rank_mat[u,] <- rank(mat[u,])
}
```


Now we fit the different correlation shrinkage algorithms (Shafer-Strimmer,
GLASSO at different scales and even CorShrink on this new data).

```{r}
cov_mat <- cov(rank_mat);
system.time(strimmer_sample <- corpcor::cov.shrink(mat))
system.time(glasso_sample_005 <- glasso::glasso(cov_mat, rho = 0.05))
system.time(glasso_sample_05 <- glasso::glasso(cov_mat, rho = 0.5))
system.time(glasso_sample_1 <- glasso::glasso(cov_mat, rho = 1))
system.time(glasso_sample_10 <- glasso::glasso(cov_mat, rho = 10))
system.time(cov_sample_ML <-  CorShrink::CorShrinkML(cov2cor(cov_mat), nsamp_mat = 550, sd_boot = FALSE, ash.control = list(mixcompdist = "normal",
                                        nullweight = 1)))

```

Now we plot the image plots of these correlation matrices.

```{r echo=FALSE, eval=TRUE}
order_index <- c();
U <- unique_tissues
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```

```{r fig.width=10, fig.height=12}
cor_result <- get(load("../output/ash_cor_only_gtex_tissues.rda"))

par(mfrow=c(3,2))
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_result[order_index,order_index, numg]),
      col=col, main=paste0("corshrink true: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)


col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov2cor(strimmer_sample))[order_index, order_index],
      col=col, main=paste0("shafer strimmer: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov2cor(glasso_sample_005$w))[order_index, order_index],
      col=col, main=paste0("glasso 0.05: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)


col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov2cor(glasso_sample_05$w))[order_index, order_index],
      col=col, main=paste0("glasso 0.5: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)


col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov_sample_ML$ash_cor_only)[order_index, order_index],
      col=col, main=paste0("corshrink fake: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)

```


## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
