---
title: "corshrink vs corpcor on soft-imputed data"
author: "Kushal K Dey"
date: "5/16/2017"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

We compare `CorShrink` with `corpcor` and `GLASSO` under a better imputation method, designed
to empower the latter methods. Instead of doing simple imputation, we use the `softImpute` method due to Mazumdar, Hastie.

```{r}
library(softImpute)
library(CorShrink)
```


## Gene 1 Example : ENSG00000115138 (POMC)

```{r}
gene <- "ENSG00000115138" ## update and rerun to see the plots for your gene
impute_method <- "svd" ## other option tested is "als"
```


```{r}
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
```

```{r}
gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))
 
person_label=read.table("../data/GTEX_V6/person_identifier_labels_with_numbers.txt");
samples_id <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,1]

samples_person <- sapply(samples_id, function(x) return(paste0(strsplit(as.character(x), "-")[[1]][1:2], collapse ="-")))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

unique_persons <- unique(samples_person)
unique_tissues <- unique(tissue_labels)
```


```{r echo=FALSE, eval=TRUE}
order_index <- c();
U <- unique_tissues
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```

```{r}
numg <- grep(gene, gene_names_1)
mat <- person_tissue_genes[,,numg]
```

We now use softImpute - 2 methods (svd and als) to get new matrices 

```{r}
matc=biScale(mat,col.scale=FALSE,row.scale=FALSE,trace=TRUE)
```

```{r}
fits3=softImpute(matc,rank.max=25,lambda=1,type=impute_method)
fitted_mat <- complete(mat,fits3,unscale=TRUE)
```

Now we apply the corpcor and GLASSO on the covariance matrix generated from this imputed 
data.

```{r}
#devtools::install_github("kkdey/CorShrink")
#library(CorShrink)
#sessionInfo()
cov_mat <- cov(fitted_mat);
system.time(strimmer_sample <- corpcor::cov.shrink(fitted_mat))
system.time(glasso_sample_005 <- glasso::glasso(cov_mat, rho = 0.05))
system.time(glasso_sample_05 <- glasso::glasso(cov_mat, rho = 0.5))
system.time(glasso_sample_1 <- glasso::glasso(cov_mat, rho = 1))
system.time(glasso_sample_10 <- glasso::glasso(cov_mat, rho = 10))
system.time(cov_sample_ML <-  CorShrinkML(cov2cor(cov_mat), nsamp_mat = 550, sd_boot = FALSE, ash.control = list(mixcompdist = "normal", nullweight = 1)))
```

```{r fig.width=10, fig.height=12}
cor_result <- get(load("../output/cor_tissues_non_ash_voom_pearson.rda"))
ash_cor_result <- get(load("../output/ash_cor_only_voom_pearson_halfuniform_gtex_tissues.rda"))
par(mfrow=c(3,2))


col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_result[order_index,order_index, numg]),
      col=col, main=paste0("sample corr with NA: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov2cor(cov_mat[order_index, order_index])),
      col=col, main=paste0("softimpute sample corr: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)



col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(ash_cor_result[order_index,order_index, numg]),
      col=col, main=paste0("corshrink true: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov2cor(strimmer_sample))[order_index, order_index],
      col=col, main=paste0("shafer strimmer: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov2cor(glasso_sample_005$w))[order_index, order_index],
      col=col, main=paste0("glasso 0.05: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)


# col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
#       rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
# image(as.matrix(cov2cor(glasso_sample_05$w))[order_index, order_index],
#       col=col, main=paste0("glasso 0.5: "), cex.main=2,
#       xaxt = "n", yaxt = "n", zlim=c(-1,1))
# axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
# axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)


col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov_sample_ML$ash_cor_only)[order_index, order_index],
      col=col, main=paste0("corshrink fake: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)

```


## Gene 1 Example : ENSG00000244734 (HBB)

```{r}
gene <- "ENSG00000244734" ## update and rerun to see the plots for your gene
impute_method <- "svd" ## other option tested is "als"
```


```{r}
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
```

```{r}
gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))
 
person_label=read.table("../data/GTEX_V6/person_identifier_labels_with_numbers.txt");
samples_id <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,1]

samples_person <- sapply(samples_id, function(x) return(paste0(strsplit(as.character(x), "-")[[1]][1:2], collapse ="-")))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

unique_persons <- unique(samples_person)
unique_tissues <- unique(tissue_labels)
```


```{r echo=FALSE, eval=TRUE}
order_index <- c();
U <- unique_tissues
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```

```{r}
numg <- grep(gene, gene_names_1)
mat <- person_tissue_genes[,,numg]
```

We now use softImpute - 2 methods (svd and als) to get new matrices 

```{r}
matc=biScale(mat,col.scale=FALSE,row.scale=FALSE,trace=TRUE)
```

```{r}
fits3=softImpute(matc,rank.max=25,lambda=1,type=impute_method)
fitted_mat <- complete(mat,fits3,unscale=TRUE)
```

Now we apply the corpcor and GLASSO on the covariance matrix generated from this imputed 
data.

```{r}
#devtools::install_github("kkdey/CorShrink")
#library(CorShrink)
#sessionInfo()
cov_mat <- cov(fitted_mat);
system.time(strimmer_sample <- corpcor::cov.shrink(fitted_mat))
system.time(glasso_sample_005 <- glasso::glasso(cov_mat, rho = 0.05))
system.time(glasso_sample_05 <- glasso::glasso(cov_mat, rho = 0.5))
system.time(glasso_sample_1 <- glasso::glasso(cov_mat, rho = 1))
system.time(glasso_sample_10 <- glasso::glasso(cov_mat, rho = 10))
system.time(cov_sample_ML <-  CorShrinkML(cov2cor(cov_mat), nsamp_mat = 550, sd_boot = FALSE, ash.control = list(mixcompdist = "normal", nullweight = 1)))
```

```{r fig.width=10, fig.height=12}
cor_result <- get(load("../output/cor_tissues_non_ash_voom_pearson.rda"))
ash_cor_result <- get(load("../output/ash_cor_only_voom_pearson_gtex_tissues.rda"))
par(mfrow=c(3,2))


col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_result[order_index,order_index, numg]),
      col=col, main=paste0("sample corr with NA: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov2cor(cov_mat[order_index, order_index])),
      col=col, main=paste0("softimpute sample corr: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)



col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(ash_cor_result[order_index,order_index, numg]),
      col=col, main=paste0("corshrink true: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov2cor(strimmer_sample))[order_index, order_index],
      col=col, main=paste0("shafer strimmer: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov2cor(glasso_sample_005$w))[order_index, order_index],
      col=col, main=paste0("glasso 0.05: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)


# col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
#       rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
# image(as.matrix(cov2cor(glasso_sample_05$w))[order_index, order_index],
#       col=col, main=paste0("glasso 0.5: "), cex.main=2,
#       xaxt = "n", yaxt = "n", zlim=c(-1,1))
# axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
# axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)


col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cov_sample_ML$ash_cor_only)[order_index, order_index],
      col=col, main=paste0("corshrink fake: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 1)

```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
