---
title: "corshrink vs correlation matrix: GTEx data"
author: "Kushal K Dey"
date: "5/12/2017"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

We propose CorShrink as a method to clean up the correlation matrix adaptively specially when the correlation values have varying degree of confidence attached 
to them as in the GTEx case, since the correlation between two tissues are based
on the number of common persons contributing both these tissues, which may vary
widely from one tissue to another.

```{r}
cor_result <- get(load("../output/cor_tissues_non_ash.rda"))
common_samples <- get(load("../output/common_samples.rda"))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]


gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))

U <- unique(tissue_labels)
```

An example correlation plot

```{r echo=FALSE, eval=TRUE}
order_index <- c();
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```


```{r}
num <- 5
plot.new()
cor_mat <- diag(1,53)+cor_result[,,num]
col=c(rev(rgb(seq(1,0,length=1000),1, seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_mat)[order_index, order_index], col=col,
      main=paste0("Corr mat: ", gene_names_1[num]), cex.main=1, xaxt = "n", yaxt = "n", zlim= c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 1)
```

Looks messy isn't it, lets see how corshrink adjusted matrix looks like.


```{r}
library(CorShrink)
source("../../CorShrink/R/CorShrinkML.R")
system.time(cor_sample_ML <-  CorShrinkML(cor_mat, nsamp_mat = as.matrix(common_samples), sd_boot = FALSE, ash.control = list(mixcompdist = "normal")))


col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_only)[order_index, order_index],
      col=col, main=paste0("CorShrink mat: ", gene_names_1[num]), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 1)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 1)

```

```{r}
num <- 5
cor_mat <- diag(1,53)+cor_result[,,num]

system.time(cor_sample_ML <-  CorShrinkML(cor_mat, common_samples, sd_boot = FALSE,
                                          ash.control = list(mixcompdist = "normal",
                                                             nullweight = 100)))

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_PD)[order_index, order_index],
      col=col, main=paste0("CorShrink mat: ", gene_names_1[num], " null: ", 100), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)

system.time(cor_sample_ML <-  CorShrinkML(cor_mat, common_samples, sd_boot = FALSE,
                                          ash.control = list(mixcompdist = "normal",
                                                             nullweight = 1000)))

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_PD)[order_index,order_index],
      col=col, main=paste0("CorShrink mat: ", gene_names_1[num], " null: ", 1000), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)

system.time(cor_sample_ML <-  CorShrinkML(cor_mat, common_samples, sd_boot = FALSE,
                                          ash.control = list(mixcompdist = "normal",
                                                             nullweight = 10)))

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_PD)[order_index,order_index],
      col=col, main=paste0("CorShrink mat: ", gene_names_1[num], " null: ", 10), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)

system.time(cor_sample_ML <-  CorShrinkML(cor_mat, common_samples, sd_boot = FALSE,
                                          ash.control = list(mixcompdist = "normal",
                                                             nullweight = 1)))

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_PD)[order_index,order_index],
      col=col, main=paste0("CorShrink mat: ", gene_names_1[num], " null: ", 1), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)


######################  different types of ash (uniform)  ###############################
num <- 5
cor_mat <- diag(1,53)+cor_result[,,num]

system.time(cor_sample_ML <-  CorShrinkML(cor_mat, common_samples, sd_boot = FALSE,
                                          ash.control = list(mixcompdist = "uniform",
                                                             nullweight = 10)))

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_PD)[order_index,order_index],
      col=col, main=paste0("CorShrink mat: ", gene_names_1[num], " comp: uniform "), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)


system.time(cor_sample_ML <-  CorShrinkML(cor_mat, common_samples, sd_boot = FALSE,
                                          ash.control = list(mixcompdist = "+uniform",
                                                             nullweight = 10)))

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_PD)[order_index,order_index],
      col=col,main=paste0("CorShrink mat: ", gene_names_1[num], " comp: +uniform "), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)


system.time(cor_sample_ML <-  CorShrinkML(cor_mat, common_samples, sd_boot = FALSE,
                                          ash.control = list(mixcompdist = "normal",
                                                             nullweight = 10)))

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_PD)[order_index,order_index],
      col=col, main=paste0("CorShrink mat: ", gene_names_1[num], " comp: normal "), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)


system.time(cor_sample_ML <-  CorShrinkML(cor_mat, common_samples, sd_boot = FALSE,
                                          ash.control = list(mixcompdist = "-uniform",
                                                             nullweight = 10)))

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_PD)[order_index,order_index],
      col=col, main=paste0("CorShrink mat: ", gene_names_1[num], " comp: -uniform "), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)


system.time(cor_sample_ML <-  CorShrinkML(cor_mat, common_samples, sd_boot = FALSE,
                                          ash.control = list(mixcompdist = "halfuniform",
                                                             nullweight = 10,
                                                             mode = 0)))

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_PD)[order_index,order_index],
      col=col, main=paste0("CorShrink mat: ", gene_names_1[num], " comp: halfuniform "), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)


system.time(cor_sample_ML <-  CorShrinkML(cor_mat, common_samples, sd_boot = FALSE,
                                          ash.control = list(mixcompdist = "normal",
                                                             nullweight = 10,
                                                             mode = "estimate")))

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(cor_sample_ML$ash_cor_PD)[order_index,order_index],
      col=col, main=paste0("CorShrink mat: shrink to est "), cex.main=1,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)
axis(2, at = seq(0, 1, length.out = ncol(cor_result[,,1])), labels = U[order_index], las=2, cex.axis = 0.4)

```


## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
