---
title: "Death time comparison study"
author: "Kushal K Dey"
date: "6/4/2017"
output: html_document
---

## Death time data reading and t-score calculation

```{r}
betahat_sebetahat <- get(load("../output/death_time_cor_nonmash.rda"))
sample_attributes <- get(load("../output/sample_attributes_filtered.rda"))
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
```

```{r}
betahat_mat <- betahat_sebetahat$betahat
sebetahat_mat <- betahat_sebetahat$sebetahat
pval_mat <- betahat_sebetahat$pval
rsquare_mat <- betahat_sebetahat$rsquare
```

```{r}
tmat <- betahat_mat/sebetahat_mat
abs_tmat <- abs(tmat)
column_sum_1 <- apply(abs_tmat, 2, function(x) return(quantile(x, 0.7)))
```

## Pathway analysis of seemingly significant genes

```{r}
names <- dimnames(person_tissue_genes)[[3]][which(column_sum_1 > 1.8)]
write.table(names, file = "../utilities/circadian_genes.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
ensembl = biomaRt::useMart("ensembl", dataset="hsapiens_gene_ensembl", host="www.ensembl.org")
res = biomaRt::getBM(attributes = c("chromosome_name", "ensembl_gene_id", "external_gene_name", "description"), filters = c("ensembl_gene_id"), values = names,  
mart = ensembl)
out <- mygene::queryMany(names,  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
cbind.data.frame(out[[4]], out[[5]])
```

```{r}
names
```


The pathway data we read.

```{r}
circadian <- read.delim("../utilities/circadian_pathways.tab")
cbind.data.frame(circadian$q.value, circadian$pathway, circadian$members_input_overlap_geneids)
```

```{r fig.height = 10}
pathway2 <- read.delim("../utilities/circadian_pathways.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 1, cexCol = 1, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Pathway analysis of genes above quantile threshold 

```{r}
column_sum_1 <- apply(abs_tmat, 2, function(x) return(quantile(x, 0.7)))
names <- dimnames(person_tissue_genes)[[3]][order(column_sum_1, decreasing = TRUE)[1:100]]
write.table(names, file = "../utilities/circadian_genes_2.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
ensembl = biomaRt::useMart("ensembl", dataset="hsapiens_gene_ensembl", host="www.ensembl.org")
res = biomaRt::getBM(attributes = c("chromosome_name", "ensembl_gene_id", "external_gene_name", "description"), filters = c("ensembl_gene_id"), values = names,  
mart = ensembl)
out <- mygene::queryMany(names,  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
```

```{r}
names
```


```{r}
circadian <- read.delim("../utilities/circadian_pathways_2.tab")
cbind.data.frame(circadian$q.value, circadian$pathway, circadian$members_input_overlap_geneids)
```

The pathway heatmap in this case 

```{r fig.height = 10}
pathway2 <- read.delim("../utilities/circadian_pathways_2.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 1, cexCol = 1, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## t-score analysis

The plots of the t-scores  for some of the genes which we know are part of the Circadian rthym.

```{r}
pathways <- read.delim("../utilities/CPDB_pathways_genes.tab")
pathway_names <- pathways[,1]
grep("Circadian", pathway_names)
```

```{r}
circadian_pathways <- pathways[grep("Circadian", pathway_names),]
gene_ids <- circadian_pathways[,4]
circadian_rhythm_genes_pre <- gene_ids[2]
circadian_clock_genes_pre <- gene_ids[3]
```

```{r}
circadian_clock_genes <- strsplit(as.character(circadian_clock_genes_pre), "[,]")[[1]]
circadian_rhythm_genes <- strsplit(as.character(circadian_rhythm_genes_pre), "[,]")[[1]]
```

### t-score plots for circadian clock genes

Plots of the t-scores for the circadian clock genes and the circadian rthym genes.

```{r}
clock_indices <- match(circadian_clock_genes, dimnames(person_tissue_genes)[[3]])
clock_indices <- clock_indices[!is.na(clock_indices)]
```

```{r}
par(mfrow=c(3,3))
for(j in 1:length(clock_indices)){
  plot(tmat[,clock_indices[j]], pch=20, cex=1, ylim=c(-3, 3),
       ylab =paste0(dimnames(person_tissue_genes)[[3]][clock_indices[j]]))
  abline(0,0)
}
```

### t-score plots for circadian rhythm genes

```{r}
rhythm_indices <- match(circadian_rhythm_genes, dimnames(person_tissue_genes)[[3]])
rhythm_indices <- rhythm_indices[!is.na(rhythm_indices)]
```

```{r}
par(mfrow=c(3,3))
for(j in 1:length(rhythm_indices)){
  plot(tmat[,rhythm_indices[j]], pch=20, cex=1, ylim=c(-3, 3),
       ylab =paste0(dimnames(person_tissue_genes)[[3]][rhythm_indices[j]]))
  abline(0,0)
}
```

### t-score plots for top quantile genes

The top t-score values we checked for out pathway analysis, we see how the scatter plots
look like for them.

```{r}
our_indices <- order(column_sum_1, decreasing = TRUE)[1:25]
dimnames(person_tissue_genes)[[3]][our_indices]
```

```{r}
par(mfrow=c(3,3))
for(j in 1:length(our_indices)){
  plot(tmat[,our_indices[j]], pch=20, cex=1, ylim=c(-3, 3),
       ylab =paste0(dimnames(person_tissue_genes)[[3]][our_indices[j]]))
  abline(0,0)
}
```

### t-score plots for random set of genes

How does a random set of 25 genes look like in its patterns across the tissues.

```{r}
random_indices <- sample(1:16069, 25, replace=FALSE)
par(mfrow=c(3,3))
for(j in 1:length(random_indices)){
  plot(tmat[,random_indices[j]], pch=20, cex=1, ylim=c(-3, 3),
       ylab =paste0(dimnames(person_tissue_genes)[[3]][random_indices[j]]))
  abline(0,0)
}
```
