---
title: "FLASH on systematic sex study"
author: "Kushal K Dey"
date: "6/14/2017"
output: html_document
---

In this script, we carry out FLASH for systematic study of sex effect on gene expression.

```{r}
library(ggplot2)
source("../code/FactorGGBar.R")
source("../code/flash_factor_postprocess.R")
```

```{r}
data <- get(load("../output/sex_cor_nonmash.rda"))
betahat <- data$betahat
sebetahat <- data$sebetahat
zdat <- betahat/sebetahat
```

```{r}
flash_out <- get(load("../output/sex_cor_flash.rda"))

common_samples <- get(load("../output/common_samples.rda"))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))


U <- unique(tissue_labels)
```

```{r}
order_index <- c();
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```


```{r fig.height=20, fig.weight =5}
loadings <- flash_out$l[order_index,]
rownames(loadings) <- 1:dim(flash_out$l)[1]
annotation <- data.frame(sample_id = 1:NROW(loadings),
                         label = factor(U[order_index], levels = U[order_index]))
FactorGGBar(loadings, annotation = annotation, panel=list(panel_rows=6))
```

The proportion of variance explained by the factors are given by 

```{r}
postprocess <- flash_factor_postprocess(flash_out$l, flash_out$f, zdat)
100*postprocess$PVE
```

It seems that the first factor corresponds to a specific effect in Breast-mammary tissue. The second factor was 
high in everything else except the breast mammary tissue which is negative. Factor 3 is strong positive in the brain tissues and also negatively related to the adipose tissues. The 4th factor is mainly expressed in the skin tissues.
It is not the case however that there are many tissue specific effects.

We look at how the factors look like? Are they sparse?

```{r}
par(mfrow=c(2,2))
plot(flash_out$f[,1], ylab = "Factor 1")
plot(flash_out$f[,2], ylab = "Factor 2")
plot(flash_out$f[,3], ylab = "Factor 3")
plot(flash_out$f[,4], ylab = "Factor 4")
```

```{r}
indices1 <- order(abs(flash_out$f[,1]), decreasing=TRUE)[1:50]
indices2 <- order(abs(flash_out$f[,2]), decreasing=TRUE)[1:50]
indices3 <- order(abs(flash_out$f[,3]), decreasing=TRUE)[1:50]
indices4 <- order(abs(flash_out$f[,4]), decreasing=TRUE)[1:50]
common_indices <- Reduce(intersect, list(indices1, indices2, indices3, indices4))
union_indices <- Reduce(union, list(indices1, indices2, indices3, indices4))
common_indices
union_indices
```

The distribution of the z-scores for some of these genes are as follows.

```{r}
par(mfrow=c(2,2))

plot(1:53, zdat[order_index,common_indices[1]], pch=20, type="b", xlab = "tissues", ylab="z scores")
text(1:53, zdat[order_index,common_indices[1]], U[order_index], cex=0.3)

plot(1:53, zdat[order_index,common_indices[2]], pch=20, type="b", xlab = "tissues", ylab="z scores")
text(1:53, zdat[order_index,common_indices[2]], U[order_index], cex=0.3)

plot(1:53, zdat[order_index,common_indices[3]], pch=20, type="b", xlab = "tissues", ylab="z scores")
text(1:53, zdat[order_index,common_indices[3]], U[order_index], cex=0.3)

```

The genes in top 25 of the first cluster 

```{r}
par(mfrow=c(3,3))
for(m in 1:50){
  plot(1:53, zdat[order_index,union_indices[m]], pch=20, type="b", xlab = "tissues", ylab="z scores")
  text(1:53, zdat[order_index,union_indices[m]], U[order_index], cex=0.3)
}
```
## Factor 3

```{r}
indices <- order(abs(flash_out$f[,3]), decreasing=TRUE)[1:50]
par(mfrow=c(3,3))
for(m in 1:50){
  plot(1:53, zdat[order_index,indices[m]], pch=20, type="b", xlab = "tissues", ylab="z scores")
  text(1:53, zdat[order_index,indices[m]], U[order_index], cex=0.3)
}
```

## Factor 4

```{r}
indices <- order(abs(flash_out$f[,4]), decreasing=TRUE)[1:50]
par(mfrow=c(3,3))
for(m in 1:50){
  plot(1:53, zdat[order_index,indices[m]], pch=20, type="b", xlab = "tissues", ylab="z scores")
  text(1:53, zdat[order_index,indices[m]], U[order_index], cex=0.3)
}
```

## Factor 5

```{r}
indices <- order(abs(flash_out$f[,5]), decreasing=TRUE)[1:50]
par(mfrow=c(3,3))
for(m in 1:50){
  plot(1:53, zdat[order_index,indices[m]], pch=20, type="b", xlab = "tissues", ylab="z scores")
  text(1:53, zdat[order_index,indices[m]], U[order_index], cex=0.3)
}
```


## Factor 8

```{r}
indices <- order(abs(flash_out$f[,8]), decreasing=TRUE)[1:50]
par(mfrow=c(3,3))
for(m in 1:50){
  plot(1:53, zdat[order_index,indices[m]], pch=20, type="b", xlab = "tissues", ylab="z scores")
  text(1:53, zdat[order_index,indices[m]], U[order_index], cex=0.3)
}
```


We see there are some genes specifically correlated to sex in the Breast mammary tissue. Also, there are some genes that show enrichment with respect to the brain tissues, although one cannot call them brain specific. Apart from the breast mammary tissue and brain tissues, the other tissues of interest are thyroid and skin.

## Enrichment of Breast Mammary tissue

```{r}
ash_data <- get(load("../output/sex_cor_ash.rda"))
zdat <- data$betahat/data$sebetahat
index <- grep("Breast", U)
zdat_thresh <- apply(abs(zdat), 2, function(x) {
            y <- median(x[index]) - median(x[-index])
            return(y)})

```

The number of genes that satisfy this are 

```{r}
imp_genes_indices <- order(zdat_thresh, decreasing = TRUE)[1:100]
imp_genes_names <- gene_names_1[imp_genes_indices]
imp_genes_names[1:10]
```

We look at the plot of the z-scores for the top 25 genes.

```{r}
par(mfrow=c(3,3))
for(m in 1:25){
  plot(1:53, zdat[order_index,imp_genes_indices[m]], pch=20, type="b", xlab = "tissues", ylab="z scores")
  text(1:53, zdat[order_index,imp_genes_indices[m]], U[order_index], cex=0.3)
}
```

We save the genes in a list and do enrichment analysis of these genes.

```{r}
write.table(imp_genes_names, file = "../utilities/sex_flash/breast_mammary.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
out <- mygene::queryMany(imp_genes_names[1:50],  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
cbind.data.frame(out$name, out$summary)
```

```{r}
pathway2 <- read.delim("../utilities/sex_flash/breast_mammary_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Enrichment in Brain tissues

```{r, echo=FALSE, eval=FALSE}
ash_data <- get(load("../output/sex_cor_ash.rda"))
lfdr_mat <- ash_data$post.lfsr
index <- grep("Brain", U)
lfdr_thresh <- apply(lfdr_mat, 2, function(x) {
            if(mean(x[index]) < 0.05 && mean(x[-index]) >  0.30){
              y <- 1
            }else{ y <- 0}
            return(y)})

```

```{r}
data <- get(load("../output/sex_cor_nonmash.rda"))
zdat <- data$betahat/data$sebetahat
index <- grep("Brain", U)
zdat_thresh <- apply(abs(zdat), 2, function(x) {
            y <- median(x[index]) - median(x[-index])
            return(y)})
```

```{r}
imp_genes_indices <- order(zdat_thresh, decreasing = TRUE)[1:100]
imp_genes_names <- gene_names_1[imp_genes_indices]
imp_genes_names[1:10]
```

```{r}
par(mfrow=c(3,3))
for(m in 1:length(imp_genes_indices)){
  plot(1:53, zdat[order_index,imp_genes_indices[m]], pch=20, type="b", xlab = "tissues", ylab="z scores")
  text(1:53, zdat[order_index,imp_genes_indices[m]], U[order_index], cex=0.3)
}
```
```{r}
write.table(imp_genes_names, file = "../utilities/sex_flash/brain.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
out <- mygene::queryMany(imp_genes_names[1:50],  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
cbind.data.frame(out$name, out$summary)
```

```{r}
pathway2 <- read.delim("../utilities/sex_flash/brain_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.1),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```


## Enrichment of skin tissues 

```{r}
data <- get(load("../output/sex_cor_nonmash.rda"))
zdat <- data$betahat/data$sebetahat
index <- grep("Skin", U)
zdat_thresh <- apply(abs(zdat), 2, function(x) {
            y <- median(x[index]) - median(x[-index])
            return(y)})
```

```{r}
imp_genes_indices <- order(zdat_thresh, decreasing = TRUE)[1:100]
imp_genes_names <- gene_names_1[imp_genes_indices]
imp_genes_names[1:10]
```

```{r}
par(mfrow=c(3,3))
for(m in 1:length(imp_genes_indices)){
  plot(1:53, zdat[order_index,imp_genes_indices[m]], pch=20, type="b", xlab = "tissues", ylab="z scores")
  text(1:53, zdat[order_index,imp_genes_indices[m]], U[order_index], cex=0.3)
}
```

```{r}
write.table(imp_genes_names, file = "../utilities/sex_flash/skin.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
out <- mygene::queryMany(imp_genes_names[1:50],  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
cbind.data.frame(out$name, out$summary)
```

```{r}
pathway2 <- read.delim("../utilities/sex_flash/skin_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.1),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Enrichment of adipose

```{r}
data <- get(load("../output/sex_cor_nonmash.rda"))
zdat <- data$betahat/data$sebetahat
index <- grep("Adipose", U)
zdat_thresh <- apply(abs(zdat), 2, function(x) {
            y <- median(x[index]) - median(x[-index])
            return(y)})
```

```{r}
imp_genes_indices <- order(zdat_thresh, decreasing = TRUE)[1:100]
imp_genes_names <- gene_names_1[imp_genes_indices]
imp_genes_names[1:10]
```

```{r}
par(mfrow=c(3,3))
for(m in 1:length(imp_genes_indices)){
  plot(1:53, zdat[order_index,imp_genes_indices[m]], pch=20, type="b", xlab = "tissues", ylab="z scores")
  text(1:53, zdat[order_index,imp_genes_indices[m]], U[order_index], cex=0.3)
}
```

```{r}
write.table(imp_genes_names, file = "../utilities/sex_flash/adipose.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
out <- mygene::queryMany(imp_genes_names[1:50],  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
cbind.data.frame(out$name, out$summary)
```

```{r}
pathway2 <- read.delim("../utilities/sex_flash/adipose_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.1),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Enrichment of Thyroid

```{r}
data <- get(load("../output/sex_cor_nonmash.rda"))
zdat <- data$betahat/data$sebetahat
index <- grep("Thyroid", U)
zdat_thresh <- apply(abs(zdat), 2, function(x) {
            y <- median(x[index]) - median(x[-index])
            return(y)})
```

```{r}
imp_genes_indices <- order(zdat_thresh, decreasing = TRUE)[1:100]
imp_genes_names <- gene_names_1[imp_genes_indices]
imp_genes_names[1:10]
```

```{r}
par(mfrow=c(3,3))
for(m in 1:length(imp_genes_indices)){
  plot(1:53, zdat[order_index,imp_genes_indices[m]], pch=20, type="b", xlab = "tissues", ylab="z scores")
  text(1:53, zdat[order_index,imp_genes_indices[m]], U[order_index], cex=0.3)
}
```

```{r}
write.table(imp_genes_names, file = "../utilities/sex_flash/thyroid.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
out <- mygene::queryMany(imp_genes_names[1:50],  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
cbind.data.frame(out$name, out$summary)
```

```{r}
pathway2 <- read.delim("../utilities/sex_flash/thyroid_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.1),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Reads expression exploration

Here we explore the log CPM expression against sex across tissues for some genes of interest that we got from above.

```{r}
betahat_sebetahat <- get(load("../output/sex_cor_ash.rda"))
sample_attributes <- get(load("../output/sample_attributes_filtered.rda"))
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
```

### ENSG00000256462 (Breast specific)

```{r}
indices <- match("ENSG00000256462", gene_names_1)
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```
### ENSG00000147889 (Breast specific)

```{r}
indices <- match("ENSG00000147889", gene_names_1)
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```

### ENSG00000259952 (Brain specific)

```{r}
indices <- match("ENSG00000259952", gene_names_1)
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```

### ENSG00000171757 (Thyroid specific)

```{r}
indices <- match("ENSG00000171757", gene_names_1)
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```

### ENSG00000236698 (Adipose specific)

```{r}
indices <- match("ENSG00000236698", gene_names_1)
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```
