---
title: 'GSEA Brain : CorShrink and CountClust'
author: "Kushal K Dey"
date: "5/12/2017"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

In this script, we try to perform some gene set enrichment analysis of the CountClust
clusters with respect to the CorShrink patterns observed.

```{r}
ash_cor_result <- get(load("../output/ash_cor_only_voom_pearson_halfuniform_gtex_tissues.rda"))
common_samples <- get(load("../output/common_samples.rda"))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

#library(data.table)
#data <- data.frame(fread("../data/GTEX_V6/cis_gene_expression.txt"))
#matdata <- t(data[,-c(1,2)])

gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))

U <- unique(tissue_labels)
```

```{r echo=FALSE, eval=TRUE}
order_index <- c();
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```

## Median comparison GSEA clusters Brain

```{r}
tab <- array(0, dim(ash_cor_result)[3])
for(m in 1:dim(ash_cor_result)[3]){
  z <- as.matrix(ash_cor_result[order_index[1:13],order_index[1:13],m])
  vec_z <- z[row(z) > col(z)]
  tab[m] <- quantile(vec_z, 0.8)
}
```

```{r}
tval <- (tab - mean(tab))/sd(tab)
names(tval) <- gene_names_1

cluster_list <- vector(mode = "list", length = 6)
for(l in 1:6){
  cluster_list[[l]] <- as.character(read.table(paste0("../utilities/gene_names_brain_clus_", l, ".txt"))[,1])[1:100]
}

brain_related_genes <- c();
for(l in 1:length(cluster_list)){
  brain_related_genes <- c(brain_related_genes, cluster_list[[l]]);
}
brain_related_genes <- unique(brain_related_genes)

other_genes <- setdiff(gene_names_1, brain_related_genes)

cluster_list[[7]] <- sample(other_genes, 100, replace = FALSE)


library(fgsea)
out <- fgsea(pathways = cluster_list,
             stats = tval,
             nperm = 100000)
out
```

The number of leadingEdge genes 

```{r}
unlist(lapply(out$leadingEdge, function(x) return(length(x))))
```

### cluster 1

```{r fig.height = 8}
genes_clus <- cluster_list[[1]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 2

```{r fig.height = 8}
genes_clus <- cluster_list[[2]][1:9]
#genes_clus <- out$leadingEdge[[2]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 3

```{r fig.height = 8}
genes_clus <- cluster_list[[3]][1:9]
#genes_clus <- out$leadingEdge[[3]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 4

```{r fig.height = 8}
genes_clus <- cluster_list[[4]][1:9]
#genes_clus <- out$leadingEdge[[4]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 5

```{r fig.height = 8}
genes_clus <- cluster_list[[5]][1:9]
#genes_clus <- out$leadingEdge[[5]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 6

```{r fig.height = 8}
genes_clus <- cluster_list[[6]][1:9]
#genes_clus <- out$leadingEdge[[6]][1:9]
matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 7

```{r fig.height = 8}
genes_clus <- cluster_list[[7]][1:9]
#genes_clus <- out$leadingEdge[[7]][1:9]
matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```


## Quantile 0.8 GSEA Brain clusters 

```{r}
tab <- array(0, dim(ash_cor_result)[3])
for(m in 1:dim(ash_cor_result)[3]){
  z <- as.matrix(ash_cor_result[order_index[1:13],order_index[1:13],m])
  vec_z <- z[row(z) > col(z)]
  tab[m] <- quantile(vec_z, 0.8)
}
```

```{r}
tval <- (tab - mean(tab))/sd(tab)
names(tval) <- gene_names_1

cluster_list <- vector(mode = "list", length = 6)
for(l in 1:6){
  cluster_list[[l]] <- as.character(read.table(paste0("../utilities/gene_names_brain_clus_", l, ".txt"))[,1])[1:100]
}

brain_related_genes <- c();
for(l in 1:length(cluster_list)){
  brain_related_genes <- c(brain_related_genes, cluster_list[[l]]);
}
brain_related_genes <- unique(brain_related_genes)

other_genes <- setdiff(gene_names_1, brain_related_genes)

cluster_list[[7]] <- sample(other_genes, 100, replace = FALSE)


library(fgsea)
out <- fgsea(pathways = cluster_list,
             stats = tval,
             nperm = 100000)
out
```

The number of leadingEdge genes 

```{r}
unlist(lapply(out$leadingEdge, function(x) return(length(x))))
```

### cluster 1

```{r fig.height = 8}
genes_clus <- cluster_list[[1]][1:9]
#genes_clus <- out$leadingEdge[[7]][1:9]
matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```
### cluster 2

```{r fig.height = 8}
genes_clus <- cluster_list[[2]][1:9]
#genes_clus <- out$leadingEdge[[2]][1:9]
matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 3

```{r fig.height = 8}
genes_clus <- cluster_list[[3]][1:9]
#genes_clus <- out$leadingEdge[[3]][1:9]
matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 4

```{r fig.height = 8}
genes_clus <- cluster_list[[4]][1:9]
#genes_clus <- out$leadingEdge[[4]][1:9]
matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 5

```{r fig.height = 8}
genes_clus <- cluster_list[[5]][1:9]
#genes_clus <- out$leadingEdge[[5]][1:9]
matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 6

```{r fig.height = 8}
genes_clus <- cluster_list[[6]][1:9]
#genes_clus <- out$leadingEdge[[6]][1:9]
matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```

### cluster 7

```{r fig.height = 8}
genes_clus <- cluster_list[[7]][1:9]
#genes_clus <- out$leadingEdge[[7]][1:9]
matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index[1:13],order_index[1:13], matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
    axis(2, at = seq(0, 1, length.out = 13), labels = U[order_index[1:13]], las=2, cex.axis = 1)
}
```


## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
