---
title: "GSEA All tissues :: CorShrink vs CountClust"
author: "Kushal K Dey"
date: "5/12/2017"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

In this script, we try to perform some gene set enrichment analysis of the CountClust
clusters with respect to the CorShrink patterns observed

```{r}
ash_cor_result <- get(load("../output/ash_cor_only_voom_pearson_halfuniform_gtex_tissues.rda"))
common_samples <- get(load("../output/common_samples.rda"))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

#library(data.table)
#data <- data.frame(fread("../data/GTEX_V6/cis_gene_expression.txt"))
#matdata <- t(data[,-c(1,2)])

gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))


U <- unique(tissue_labels)
```

```{r echo=FALSE, eval=TRUE}
order_index <- c();
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```

## Brain related 

```{r}
#####################  Brain high  ############################

tab <- array(0, dim(ash_cor_result)[3])
for(m in 1:dim(ash_cor_result)[3]){
  temp <- ash_cor_result[order_index[1:13], order_index[(1:13)],m]
  tab[m] <-  median(temp[row(temp) > col(temp)])
}

tval <- (tab - mean(tab))/sd(tab)
names(tval) <- gene_names_1
```

```{r}
cluster_list <- vector(mode = "list", length = 20)
for(l in 1:20){
  cluster_list[[l]] <- as.character(read.table(paste0("../utilities/gene_names_clus_", l, ".txt"))[,1])
}

related_genes <- c();
for(l in 1:length(cluster_list)){
  related_genes <- c(related_genes, cluster_list[[l]]);
}
related_genes <- unique(related_genes)

other_genes <- setdiff(gene_names_1, related_genes)

cluster_list[[7]] <- sample(other_genes, 100, replace = FALSE)


library(fgsea)
out <- fgsea(pathways = cluster_list,
             stats = tval,
             nperm = 10000)

out
```
It is dominated by the Clusters 17 and 19 which are blood related tissues in terms of positive effects. We need to
carry out more focussed analysis.


## High intra corr in Brain, low intra corr in others

```{r}
#####################  Brain high, others low  ############################

tab <- array(0, dim(ash_cor_result)[3])
for(m in 1:dim(ash_cor_result)[3]){
  temp <- ash_cor_result[order_index[1:13], order_index[(1:13)],m]
  temp1 <- ash_cor_result[order_index[-(1:13)], order_index[-(1:13)],m]
  tab[m] <- median(temp[row(temp) > col(temp)]) - median(temp1[row(temp1) > col(temp1)])
}

tval <- (tab - mean(tab))/sd(tab)
names(tval) <- gene_names_1
```

```{r}
cluster_list <- vector(mode = "list", length = 20)
for(l in 1:20){
  cluster_list[[l]] <- as.character(read.table(paste0("../utilities/gene_names_clus_", l, ".txt"))[,1])
}

related_genes <- c();
for(l in 1:length(cluster_list)){
  related_genes <- c(related_genes, cluster_list[[l]]);
}
related_genes <- unique(related_genes)

other_genes <- setdiff(gene_names_1, related_genes)

cluster_list[[7]] <- sample(other_genes, 100, replace = FALSE)


library(fgsea)
out <- fgsea(pathways = cluster_list,
             stats = tval,
             nperm = 10000)

out
```


```{r}
unlist(lapply(out$leadingEdge, function(x) return(length(x))))
```

### cluster 3

```{r fig.height = 8}
genes_clus <- cluster_list[[3]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index,order_index, matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
    axis(2, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
}
```

### cluster 9

```{r fig.height = 8}
genes_clus <- cluster_list[[9]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index,order_index, matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
    axis(2, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
}
```


### cluster 15

```{r fig.height = 8}
genes_clus <- cluster_list[[15]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index,order_index, matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
    axis(2, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
}
```


### cluster 16

```{r fig.height = 8}
genes_clus <- cluster_list[[16]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index,order_index, matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
    axis(2, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
}
```


### cluster 17

```{r fig.height = 8}
genes_clus <- cluster_list[[17]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index,order_index, matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
    axis(2, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
}
```

Which tissues do these clusters corrsspond to?

clusters 15, 16 mainly correspond to Esophagus Mucosa and Lung. Cluster 3 corresponds to Breast and Adipose - the 
fatty issues. Cluster 9, which is one of the rare clusters with strong negative effect in gsea corresponds to the
Brain-spinal cord genes, which we know have less correlation among the brain tissues.

## High Brain vs other tissues 

```{r}
tab <- array(0, dim(ash_cor_result)[3])
for(m in 1:dim(ash_cor_result)[3]){
  temp <- ash_cor_result[order_index[1:13], order_index[-(1:13)],m]
  tab[m] <- median(temp[row(temp) > col(temp)])
}

tval <- (tab - mean(tab))/sd(tab)
names(tval) <- gene_names_1
```

```{r}
cluster_list <- vector(mode = "list", length = 20)
for(l in 1:20){
  cluster_list[[l]] <- as.character(read.table(paste0("../utilities/gene_names_clus_", l, ".txt"))[,1])
}

related_genes <- c();
for(l in 1:length(cluster_list)){
  related_genes <- c(related_genes, cluster_list[[l]]);
}
related_genes <- unique(related_genes)

other_genes <- setdiff(gene_names_1, related_genes)

cluster_list[[7]] <- sample(other_genes, 100, replace = FALSE)


library(fgsea)
out <- fgsea(pathways = cluster_list,
             stats = tval,
             nperm = 10000)

out
```
The ones with high positive effects are Clusters 17 and 19, which are both shared by blood. 
There are several clusters with high negative effects such as Cluster 14 (Heart), clusters 2 and 9 (Brain
specific), Cluster 4 (Colon, Cervix etc), Cluster 5 (Artery) and Cluster 6 (Skin), which is a wide raneg of
tissue specific genes. Blood seems to be the exception, which sort of makes sense, because blood may be contaminated
with brain tissues.


### cluster 17

```{r fig.height = 8}
genes_clus <- cluster_list[[17]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index,order_index, matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
    axis(2, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
}
```

### cluster 19

```{r fig.height = 8}
genes_clus <- cluster_list[[19]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index,order_index, matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
    axis(2, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
}
```


### cluster 4

```{r fig.height = 8}
genes_clus <- cluster_list[[4]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index,order_index, matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
    axis(2, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
}
```


## Heart related


```{r}
tab <- array(0, dim(ash_cor_result)[3])
for(m in 1:dim(ash_cor_result)[3]){
  tab[m] <- ash_cor_result[32, 33 ,m]
}

tval <- (tab - mean(tab))/sd(tab)
names(tval) <- gene_names_1
```

```{r}
cluster_list <- vector(mode = "list", length = 20)
for(l in 1:20){
  cluster_list[[l]] <- as.character(read.table(paste0("../utilities/gene_names_clus_", l, ".txt"))[,1])
}

related_genes <- c();
for(l in 1:length(cluster_list)){
  related_genes <- c(related_genes, cluster_list[[l]]);
}
related_genes <- unique(related_genes)

other_genes <- setdiff(gene_names_1, related_genes)

cluster_list[[7]] <- sample(other_genes, 100, replace = FALSE)


library(fgsea)
out <- fgsea(pathways = cluster_list,
             stats = tval,
             nperm = 10000)

out
```


### cluster 1

```{r fig.height = 8}
genes_clus <- cluster_list[[1]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index,order_index, matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
    axis(2, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
}
```

### cluster 14

```{r fig.height = 8}
genes_clus <- cluster_list[[2]][1:9]
#genes_clus <- out$leadingEdge[[1]][1:9]

matched_indices <- match(genes_clus, gene_names_1)
par(mfrow = c(3,3))
for(m in 1:length(matched_indices)){
      col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
          rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
    image(as.matrix(ash_cor_result[order_index,order_index, matched_indices[m]]),
          col=col, main=paste0("corshrink:", genes_clus[m]), cex.main=2,
          xaxt = "n", yaxt = "n", zlim=c(-1,1))
    axis(1, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
    axis(2, at = seq(0, 1, length.out = 53), labels = U, las=2, cex.axis = 0.5)
}
```

## Skin related

```{r}
tab <- array(0, dim(ash_cor_result)[3])
for(m in 1:dim(ash_cor_result)[3]){
  tab[m] <- ash_cor_result[44, 45, m]
}

tval <- (tab - mean(tab))/sd(tab)
names(tval) <- gene_names_1
```

```{r}
cluster_list <- vector(mode = "list", length = 20)
for(l in 1:20){
  cluster_list[[l]] <- as.character(read.table(paste0("../utilities/gene_names_clus_", l, ".txt"))[,1])
}

related_genes <- c();
for(l in 1:length(cluster_list)){
  related_genes <- c(related_genes, cluster_list[[l]]);
}
related_genes <- unique(related_genes)

other_genes <- setdiff(gene_names_1, related_genes)

cluster_list[[7]] <- sample(other_genes, 100, replace = FALSE)


library(fgsea)
out <- fgsea(pathways = cluster_list,
             stats = tval,
             nperm = 10000)

out
```


## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
