---
title: "MASH on systemic sex study (canonical)"
author: "Kushal K Dey"
date: "6/17/2017"
output: html_document
---
```{r}
library(mashr)
```

```{r}
data <- get(load("../output/sex_cor_nonmash.rda"))
betahat <- data$betahat
sebetahat <- data$sebetahat
zdat <- betahat/sebetahat
```

```{r}
common_samples <- get(load("../output/common_samples.rda"))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))

U <- unique(tissue_labels)
```


```{r}
mash_dat <- get(load("../output/mash_sex_canonical.rda"))
```

```{r}
grep("Breast", U)
```

```{r}
barplot(get_estimated_pi(mash_dat), las = 2, cex.names = 0.3)
```

```{r}
U[c(21, 38)]
```

I feel the brain enrichment is getting merged with the simple het matrices 1 and 2 (corresponding to 0.25 and 0.5), which is why we do not see the brain enrichment. But the great thing is we see the singleton 21 corresponding to Breast Mammary tissue to be high.

## Breast specific genes 

```{r}
breast_specific <- apply(mash_dat$result$lfsr, 1, function(x) {
      if(x[21] < 0.05 && min(x[-21])> 0.9){
        return(1)
      }else{
        return(0)
      }
})
```

```{r}
length(which(breast_specific==1))
```

We look at the gene set enrichment analysis of these genes 

```{r}
imp_genes <- gene_names_1[which(breast_specific==1)]
imp_genes[1:10]
```

```{r}
write.table(imp_genes, file = "../utilities/sex_mash/breast.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
pathway2 <- read.delim("../utilities/sex_mash/breast_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Shared genes 

```{r}
num_lfsr_below_thresh <- apply(mash_dat$result$lfsr, 1, function(x) return(length(which(x < 0.05))))

imp_indices <- which(num_lfsr_below_thresh == 53)
imp_genes <- gene_names_1[imp_indices]
length(imp_genes)
imp_genes[1:10]
```

```{r}
write.table(imp_genes, file = "../utilities/sex_mash/all.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
pathway2 <- read.delim("../utilities/sex_mash/all_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

```{r}
num_lfsr_below_thresh <- apply(mash_dat$result$lfsr, 1, function(x) return(length(which(x < 0.05))))

imp_indices <- which(num_lfsr_below_thresh > 25)
imp_genes <- gene_names_1[imp_indices]
length(imp_genes)
imp_genes[1:10]
```

```{r}
write.table(imp_genes, file = "../utilities/sex_mash/half.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
pathway2 <- read.delim("../utilities/sex_mash/half_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

```{r}
num_lfsr_below_thresh <- apply(mash_dat$result$lfsr, 1, function(x) return(length(which(x < 0.05))))

imp_indices <- which(num_lfsr_below_thresh > 35)
imp_genes <- gene_names_1[imp_indices]
length(imp_genes)
imp_genes[1:10]
```

```{r}
write.table(imp_genes, file = "../utilities/sex_mash/three_quarter.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
pathway2 <- read.delim("../utilities/sex_mash/three_quarter_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Brain specific genes 

```{r}
grep("Brain", U)
```

```{r}
brain_specific <- apply(mash_dat$result$lfsr, 1, function(x) 
  {
    if(min(x[grep("Brain", U)]) < 0.05 && median(x[-grep("Brain", U)]) > 0.3){
      return(1)
    }else{
      return(0)
    }
  })
imp_genes <- gene_names_1[which(brain_specific==1)]
length(imp_genes)
```

```{r}
write.table(imp_genes, file = "../utilities/sex_mash/brain.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
pathway2 <- read.delim("../utilities/sex_mash/brain_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Muscle specific genes 

```{r}
grep("Muscle", U)
```

```{r}
muscle_specific <- apply(mash_dat$result$lfsr, 1, function(x) 
  {
    if(min(x[grep("Muscle", U)]) < 0.05 && median(x[-grep("Muscle", U)]) > 0.5){
      return(1)
    }else{
      return(0)
    }
  })
imp_genes <- gene_names_1[which(muscle_specific==1)]
length(imp_genes)
```

```{r}
imp_genes[1:10]
```

```{r}
write.table(imp_genes, file = "../utilities/sex_mash/muscle.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
pathway2 <- read.delim("../utilities/sex_mash/muscle_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```


## Reads expression profiling 

Here we do an expression profiling of some of the important genes we obtained from above

### ENSG00000136854

```{r}
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
sample_attributes <- get(load("../output/sample_attributes_filtered.rda"))
```

```{r}
indices <- match("ENSG00000149531", gene_names_1)
```

```{r}
betahat_sebetahat_2 <- get(load("../output/sex_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
plot(1:53, tmat[,indices], type = "b", xlab = "tissues", ylab="t scores")
text(1:53, tmat[,indices], U, cex=0.3)
```


```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```


### ENSG00000000457

```{r}
indices <- match("ENSG00000000457", gene_names_1)
```

```{r}
betahat_sebetahat_2 <- get(load("../output/sex_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
plot(1:53, tmat[,indices], type = "b", xlab = "tissues", ylab="t scores")
text(1:53, tmat[,indices], U, cex=0.3)
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```



### ENSG00000080007

```{r}
indices <- match("ENSG00000080007", gene_names_1)
```

```{r}
betahat_sebetahat_2 <- get(load("../output/sex_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
plot(1:53, tmat[,indices], type = "b", xlab = "tissues", ylab="t scores")
text(1:53, tmat[,indices], U, cex=0.3)
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```

### ENSG00000120688

```{r}
indices <- match("ENSG00000120688", gene_names_1)
```

```{r}
betahat_sebetahat_2 <- get(load("../output/sex_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
plot(1:53, tmat[,indices], type = "b", xlab = "tissues", ylab="t scores")
text(1:53, tmat[,indices], U, cex=0.3)
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```
