---
title: "MASH on systemic circadian study (canonical)"
author: "Kushal K Dey"
date: "6/18/2017"
output: html_document
---

```{r}
library(mashr)
```

```{r}
data <- get(load("../output/death_time_cor_nonmash.rda"))
betahat <- data$betahat
sebetahat <- data$sebetahat
zdat <- betahat/sebetahat
```

```{r}
common_samples <- get(load("../output/common_samples.rda"))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))

U <- unique(tissue_labels)
```


```{r}
mash_dat <- get(load("../output/mash_circadian_canonical.rda"))
```

```{r}
grep("Breast", U)
```

```{r}
barplot(get_estimated_pi(mash_dat), las = 2, cex.names = 0.3)
```

```{r}
U[c(7, 25, 31, 33, 35)]
```

Bladder, Cervix and Fallopian Tube show up because of having very small number of samples in them.

The interesting two tissues that show up are the Heart, Liver and Brain is clubbed into the equal effects and the het matrices.

## Liver specific genes 

```{r}
liver_specific <- apply(mash_dat$result$lfsr, 1, function(x) {
      if(x[35] < 0.05 && min(x[-35])> 0.9){
        return(1)
      }else{
        return(0)
      }
})
```

```{r}
length(which(liver_specific==1))
```

We look at the gene set enrichment analysis of these genes 

```{r}
imp_genes <- gene_names_1[which(liver_specific==1)]
imp_genes[1:10]
```

```{r}
write.table(imp_genes, file = "../utilities/circadian_mash/liver.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
pathway2 <- read.delim("../utilities/circadian_mash/liver_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Heart Left Ventricle specific genes 

```{r}
heart_specific <- apply(mash_dat$result$lfsr, 1, function(x) {
      if(x[33] < 0.1 && median(x[-33])> 0.5){
        return(1)
      }else{
        return(0)
      }
})
```

```{r}
length(which(heart_specific==1))
```

We look at the gene set enrichment analysis of these genes 

```{r}
imp_genes <- gene_names_1[which(heart_specific==1)]
imp_genes[1:10]
```

```{r}
write.table(imp_genes, file = "../utilities/circadian_mash/heart.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

No enriched pathways

## Shared genes 

```{r}
num_lfsr_below_thresh <- apply(mash_dat$result$lfdr, 1, function(x) return(length(which(x < 0.05))))

imp_indices <- which(num_lfsr_below_thresh == 53)
imp_genes <- gene_names_1[imp_indices]
length(imp_genes)
imp_genes[1:10]
```

```{r}
write.table(imp_genes, file = "../utilities/circadian_mash/all.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
pathway2 <- read.delim("../utilities/circadian_mash/all_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

```{r}
num_lfsr_below_thresh <- apply(mash_dat$result$lfdr, 1, function(x) return(length(which(x < 0.05))))

imp_indices <- which(num_lfsr_below_thresh > 25)
imp_genes <- gene_names_1[imp_indices]
length(imp_genes)
imp_genes[1:10]
```

```{r}
write.table(imp_genes, file = "../utilities/circadian_mash/half.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
pathway2 <- read.delim("../utilities/circadian_mash/half_pathway.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Brain specific genes 

```{r}
grep("Brain", U)
```

```{r}
brain_specific <- apply(mash_dat$result$lfsr, 1, function(x) 
  {
    if(min(x[grep("Brain", U)]) < 0.1 && median(x[-grep("Brain", U)]) > 0.2){
      return(1)
    }else{
      return(0)
    }
  })
imp_genes <- gene_names_1[which(brain_specific==1)]
length(imp_genes)
imp_genes
```

```{r}
write.table(imp_genes, file = "../utilities/circadian_mash/brain.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

mRNA pathway is the only enriched pathway.

## Reads expression profiling 

Here we do an expression profiling of some of the important genes we obtained from above

```{r}
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
sample_attributes <- get(load("../output/sample_attributes_filtered.rda"))
```

```{r}
times <- strsplit(as.character(sample_attributes$DTHTIME), "[:]")
time_stamp <- unlist(lapply(times, function(x){
  y <- as.numeric(x[1])
  z <- as.numeric(x[2])
  w <- y*60+z
  return(w)
}))/(24*60)

cos_times <- cos(2*pi*time_stamp)
uncos_times <- 2*pi*time_stamp
na_indices <- which(is.na(cos_times))

person_tissue_genes_1 <- person_tissue_genes[-na_indices,,]
cos_times_1 <- cos_times[-na_indices]
uncos_times_1 <- uncos_times[-na_indices]
```


### ENSG00000005436

```{r}
indices <- match("ENSG00000005436", gene_names_1)
```

```{r}
betahat_sebetahat_2 <- get(load("../output/death_time_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
plot(1:53, tmat[,indices], type = "b", xlab = "tissues", ylab="t scores")
text(1:53, tmat[,indices], U, cex=0.3)
```

```{r}
par(mfrow=c(3,3))
for(j in 1:53){
  plot(uncos_times_1, person_tissue_genes_1[,j, indices],
       ylab = dimnames(person_tissue_genes)[[2]][j])
  #lines(loess.smooth(uncos_times_1,person_tissue_genes_1[,j, indices[1]]), 
  #      col="red", lty=2, lwd=2)
}
```

### ENSG00000006194

```{r}
indices <- match("ENSG00000006194", gene_names_1)
```

```{r}
betahat_sebetahat_2 <- get(load("../output/death_time_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
plot(1:53, tmat[,indices], type = "b", xlab = "tissues", ylab="t scores")
text(1:53, tmat[,indices], U, cex=0.3)
```

```{r}
par(mfrow=c(3,3))
for(j in 1:53){
  plot(uncos_times_1, person_tissue_genes_1[,j, indices],
       ylab = dimnames(person_tissue_genes)[[2]][j])
  #lines(loess.smooth(uncos_times_1,person_tissue_genes_1[,j, indices[1]]), 
  #      col="red", lty=2, lwd=2)
}
```

### ENSG00000100029

```{r}
indices <- match("ENSG00000100029", gene_names_1)
```

```{r}
betahat_sebetahat_2 <- get(load("../output/death_time_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
plot(1:53, tmat[,indices], type = "b", xlab = "tissues", ylab="t scores")
text(1:53, tmat[,indices], U, cex=0.3)
```

```{r}
par(mfrow=c(3,3))
for(j in 1:53){
  plot(uncos_times_1, person_tissue_genes_1[,j, indices],
       ylab = dimnames(person_tissue_genes)[[2]][j])
  #lines(loess.smooth(uncos_times_1,person_tissue_genes_1[,j, indices[1]]), 
  #      col="red", lty=2, lwd=2)
}
```

### ENSG00000111737

```{r}
indices <- match("ENSG00000111737", gene_names_1)
```

```{r}
betahat_sebetahat_2 <- get(load("../output/death_time_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
plot(1:53, tmat[,indices], type = "b", xlab = "tissues", ylab="t scores")
text(1:53, tmat[,indices], U, cex=0.3)
```

```{r}
par(mfrow=c(3,3))
for(j in 1:53){
  plot(uncos_times_1, person_tissue_genes_1[,j, indices],
       ylab = dimnames(person_tissue_genes)[[2]][j])
  #lines(loess.smooth(uncos_times_1,person_tissue_genes_1[,j, indices[1]]), 
  #      col="red", lty=2, lwd=2)
}
```

### ENSG00000004660 (all)

```{r}
indices <- match("ENSG00000004660", gene_names_1)
```

```{r}
betahat_sebetahat_2 <- get(load("../output/death_time_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
plot(1:53, tmat[,indices], type = "b", xlab = "tissues", ylab="t scores")
text(1:53, tmat[,indices], U, cex=0.3)
```

```{r}
par(mfrow=c(3,3))
for(j in 1:53){
  plot(uncos_times_1, person_tissue_genes_1[,j, indices],
       ylab = dimnames(person_tissue_genes)[[2]][j])
  #lines(loess.smooth(uncos_times_1,person_tissue_genes_1[,j, indices[1]]), 
  #      col="red", lty=2, lwd=2)
}
```

### ENSG00000125510 (brain)

```{r}
indices <- match("ENSG00000125510", gene_names_1)
```

```{r}
betahat_sebetahat_2 <- get(load("../output/death_time_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
plot(1:53, tmat[,indices], type = "b", xlab = "tissues", ylab="t scores")
text(1:53, tmat[,indices], U, cex=0.3)
```

```{r}
par(mfrow=c(3,3))
for(j in 1:53){
  plot(uncos_times_1, person_tissue_genes_1[,j, indices],
       ylab = dimnames(person_tissue_genes)[[2]][j])
  #lines(loess.smooth(uncos_times_1,person_tissue_genes_1[,j, indices[1]]), 
  #      col="red", lty=2, lwd=2)
}
```


