---
title: "Sex comparison with ash"
author: "Kushal K Dey"
date: "6/8/2017"
output: html_document
---

In this script, we look at the SEX comparison problem in the context of using lfsr for the adaptive shrinkage on the tissues for each gene. Adaptive shrinkage is performed using a normal distribution centered about 0. 

```{r}
ash_cor_mat <- get(load("../output/ash_cor_only_voom_pearson_halfuniform_gtex_tissues.rda"))
```

```{r}
common_samples <- get(load("../output/common_samples.rda"))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

#library(data.table)
#data <- data.frame(fread("../data/GTEX_V6/cis_gene_expression.txt"))
#matdata <- t(data[,-c(1,2)])

gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))


U <- unique(tissue_labels)
```

```{r echo=FALSE, eval=TRUE}
order_index <- c();
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```


```{r}
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
betahat_sebetahat <- get(load("../output/sex_cor_ash.rda"))
sample_attributes <- get(load("../output/sample_attributes_filtered.rda"))
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
```

```{r}
betahat_mat <- betahat_sebetahat$post.mean
sebetahat_mat <- betahat_sebetahat$post.sd
lfsr_mat <- betahat_sebetahat$post.lfsr
```

##  lfsr threshold 0.05

We look at which genes have the most number of lfsr values below a chosen threshold.

```{r}
num_lfsr_below_thresh <- apply(lfsr_mat, 2, function(x) return(length(which(x < 0.05))))

imp_genes <- order(num_lfsr_below_thresh, decreasing = TRUE)[1:10]
num_lfsr_below_thresh[imp_genes]

names <- colnames(betahat_mat)[imp_genes]
```

```{r}
write.table(names, file = "../utilities/sex_ash/sex_genes_ash_0_05.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
out <- mygene::queryMany(names,  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
cbind.data.frame(out$name, out$summary)
```

```{r}
pathway2 <- read.delim("../utilities/sex_ash/sex_genes_ash_0_05_pathways_1.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 1, cexCol = 1, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

```{r}
num_lfsr_below_thresh <- apply(lfsr_mat, 2, function(x) return(length(which(x < 0.05))))

imp_genes <- order(num_lfsr_below_thresh, decreasing = TRUE)[1:25]
num_lfsr_below_thresh[imp_genes]

names <- colnames(betahat_mat)[imp_genes]
```

```{r}
write.table(names, file = "../utilities/sex_ash/sex_genes_ash_0_05_2.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
out <- mygene::queryMany(names,  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
cbind.data.frame(out$name, out$summary)
```

```{r}
pathway2 <- read.delim("../utilities/sex_ash/sex_genes_ash_0_05_pathways_2.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 1, cexCol = 1, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

##  lfsr threshold 0.1

We look at which genes have the most number of lfsr values below a chosen threshold.

```{r}
num_lfsr_below_thresh <- apply(lfsr_mat, 2, function(x) return(length(which(x < 0.1))))

imp_genes <- order(num_lfsr_below_thresh, decreasing = TRUE)[1:25]
num_lfsr_below_thresh[imp_genes]

names <- colnames(betahat_mat)[imp_genes]
```

```{r}
write.table(names, file = "../utilities/sex_ash/sex_genes_ash_0_1.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
out <- mygene::queryMany(names,  scopes="ensembl.gene", fields=c("name", "summary"), species="human");
cbind.data.frame(out$name, out$summary)
```

```{r}
pathway2 <- read.delim("../utilities/sex_ash/sex_genes_ash_0_1_pathways.tab")
if(length(which(pathway2$q.value < 0.05)) > 2){
  pathway2 <- pathway2[which(pathway2$q.value < 0.05),]
}
maxpath = 75
pathway <- pathway2[1:min(maxpath, dim(pathway2)[1]), ]
ensembl_genes <- pathway$members_input_overlap
pathway_list_genes <- sapply(ensembl_genes, function(x) return(strsplit(as.character(x), "; ")))
all_genes <- Reduce(union, pathway_list_genes)
pathway_mat <- matrix(0, length(pathway_list_genes), length(all_genes))
for(l in 1:dim(pathway_mat)[1]){
  pathway_mat[l, match(pathway_list_genes[[l]], all_genes)] <- 1
}

pathway_bimat <- (pathway_mat)%*%t(pathway_mat)
pathway_bimat_prop <- diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1]) %*% pathway_bimat %*% diag(1/sqrt(diag(pathway_bimat)), dim(pathway_bimat)[1])

col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
heatmap(pathway_bimat_prop, labRow = pathway$pathway, labCol = pathway$pathway, cexRow = 0.3, cexCol = 0.3, col = col, zlim = c(-1,1),
        scale = "none", margins = c(5, 5))
```

## Review of genes 

A sex based study across tissues for genes (that misses pseudogenes) is in the following
[Article](http://biorxiv.org/content/biorxiv/early/2016/10/20/082289.full.pdf).

The important genes are **ENSG00000225490**, **ENSG00000149531**,  **ENSG00000236698**, **ENSG00000258484**,  **ENSG00000080007**  and **ENSG00000180574**.

We know very little about ENSG00000225490 as it does not appear to be a valid gene. I mainly found this [Article](http://www.bcgsc.ca/people/malachig/htdocs/alexa_platform/alexa_seq/Oligo/genes/chr20_2/ENSG00000225490.htm).

FRG1BP or ENSG00000149531 is another pseudoegene. 

ENSG00000236698 or EIF1AXP1 is X-linked pseudogene. [Article](http://www.bcgsc.ca/people/malachig/htdocs/alexa_platform/alexa_seq/AllenBrain/genes/chr1_2/ENSG00000236698.htm).

SPESP1 (sperm equatorial segment protein 1) or ENSG00000258484 is a gene enriched in Testis [Page](http://www.proteinatlas.org/ENSG00000258484-SPESP1/tissue).

[Article](http://biorxiv.org/content/biorxiv/early/2016/10/20/082289.full.pdf) looks into some of these genes to have shared sexual effects across tissues.

DDX43 or ENSG00000080007 is highly expressed in human tissues.

ENSG00000180574 or AC068775.1 is an eukaryotic translation initiation factor.

## ENSG00000225490

Check the read expression profile in GTEx here 

[ENSG00000225490](https://gtexportal.org/home/gene/ENSG00000225490)

```{r}
betahat_sebetahat_2 <- get(load("../output/sex_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
```

```{r}
indices <- match("ENSG00000225490", colnames(betahat_mat))
```


```{r}
plot(tmat[, indices], col="red")
abline(0,0)
```

Which are the tissues with lowest lfsr 

```{r}
sort(lfsr_mat[,indices], decreasing = FALSE)[1:5]
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```

```{r}
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(ash_cor_mat[order_index,order_index, indices]),
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
```

## ENSG00000149531

Check the read expression profile in GTEx here 

[ENSG00000149531](https://gtexportal.org/home/gene/ENSG00000149531)

```{r}
betahat_sebetahat_2 <- get(load("../output/sex_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
```

```{r}
indices <- match("ENSG00000149531", colnames(betahat_mat))
```


```{r}
plot(tmat[, indices], col="red")
abline(0,0)
```

Which are the tissues with lowest lfsr 

```{r}
sort(lfsr_mat[,indices], decreasing = FALSE)[1:5]
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```

```{r}
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(ash_cor_mat[order_index,order_index, indices]),
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
```

## ENSG00000236698 

Check the read expression profile in GTEx here 

[ENSG00000236698](https://gtexportal.org/home/gene/ENSG00000236698)

```{r}
betahat_sebetahat_2 <- get(load("../output/sex_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
```

```{r}
indices <- match("ENSG00000236698", colnames(betahat_mat))
```


```{r}
plot(tmat[, indices], col="red")
abline(0,0)
```

Which are the tissues with lowest lfsr 

```{r}
sort(lfsr_mat[,indices], decreasing = FALSE)[1:5]
```


```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```

```{r}
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(ash_cor_mat[order_index,order_index, indices]),
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
```


## ENSG00000258484 or SPESP1

Check the read expression profile in GTEx here 

[ENSG00000258484](https://gtexportal.org/home/gene/ENSG00000258484)

```{r}
betahat_sebetahat_2 <- get(load("../output/sex_cor_nonmash.rda"))
tmat <- betahat_sebetahat_2$betahat/betahat_sebetahat_2$sebetahat
```

```{r}
indices <- match("ENSG00000258484", colnames(betahat_mat))
```


```{r}
plot(tmat[, indices], col="red")
abline(0,0)
```

Which are the tissues with lowest lfsr 

```{r}
sort(lfsr_mat[,indices], decreasing = FALSE)[1:5]
```

```{r}
par(mfrow = c(3,3))
for(j in 1:53){
  ll <-lm(person_tissue_genes[,j,indices] ~ sample_attributes$SEX)
  temp1 <- person_tissue_genes[,j,indices]
  temp2 <- sample_attributes$SEX
  temp11 <- temp1[which(!is.na(temp1))]
  temp21 <- temp2[which(!is.na(temp1))]
  temp1_M  <- temp11[which(temp21 == 1)]
  temp1_F  <- temp11[which(temp21 == 2)]
  plot(temp11, col=factor(temp21), ylab = dimnames(person_tissue_genes)[[2]][j])
}
```

```{r}
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(ash_cor_mat[order_index,order_index, indices]),
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = U[order_index], las=2, cex.axis = 0.5)
```



