---
title: "Spcov vs CorShrink"
author: "Kushal K Dey"
date: "5/24/2017"
output: html_document
---

In this script, we compare between the spcov method due to Bien and Tibshirani, under
different proposals, with our CorShrink approach.

## Data preparation

```{r}
library(spcov)
```

```{r}
gene <- "ENSG00000115138" ## update and rerun to see the plots for your gene
```

The number of samples per tissue 

```{r}
common_samples <- get(load("../output/common_samples.rda"))
```

```{r}
gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))
 
person_label=read.table("../data/GTEX_V6/person_identifier_labels_with_numbers.txt");
samples_id <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,1]

samples_person <- sapply(samples_id, function(x) return(paste0(strsplit(as.character(x), "-")[[1]][1:2], collapse ="-")))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

unique_persons <- unique(samples_person)
unique_tissues <- unique(tissue_labels)
```

```{r echo=FALSE, eval=TRUE}
order_index <- c();
U <- unique_tissues
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```

We load the sample correlation matrix and the CorShrink matrix.

## Sample correlation

```{r fig.width=10, fig.height=12}
cor_result <- get(load("../output/cor_tissues_non_ash_voom_pearson.rda"))
cor_result[is.na(cor_result)] = 0
ash_cor_result <- get(load("../output/ash_cor_only_voom_pearson_normal_gtex_tissues.rda"))
```

```{r}
numg <- match(gene, gene_names_1)
cor_mat <- diag(1, 53) + cor_result[,,numg]
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(cor_mat[order_index, order_index],
      col=col, main=paste0("sample corr: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

## CorShrink matrix

```{r}
numg <- match(gene, gene_names_1)
ash_cor_mat <- ash_cor_result[,,numg]
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(ash_cor_mat)[order_index, order_index],
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)

```

We now apply the spcov method on the correlation matrix with P by inverse of the 
number of common samples.

Now the `cor_mat` matrix is not a PD matrix, which you can check from below 

```{r}
eig <- eigen(cor_mat)
eig$values
```

Note that this is because the correlations in this matrix have not been 
calculated on the same set of observations.

Can **spcov** still work despite it not being a valid covariance matrix?

```{r}
cor_mat_pd <- Matrix::nearPD(cor_mat, corr=TRUE)
cor_mat_pd <- as.matrix(cor_mat_pd$mat)
```


## spcov with P = 1/ (common_samples)


```{r}
P <- 1/common_samples
P[P==Inf] = 1
diag(P) <- 0
P[1:5,1:5]
```

```{r}
lam <- 0.06
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

```{r}
lam <- 0.1
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

```{r}
lam <- 1
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```
```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```
```{r}
lam <- 10
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```
```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

Problems with spcov 

- You have to choose the $\lambda$ well. Here we choose the one from common samples.
  That automatically takes the separate scaling for separate variable pair corr
  into account. That seems like the best way to tackle the correlation issue.
  
- The shrinkage does not seem too strong for smaller values of $\lambda$ and
  also for high $\lambda$, the shrinkage is strong throughout the correlation
  matrix.
  
- However, the feeling is CorShrink does a better job at separately shrinking
  the correlations. For instance, the negative correlation values still remain
  even after very strong shrinkage. Also it is not a good thing that for large 
  $\lambda$, there is strong shrinkage even in the brain tissues.
  
We try to elevate the differential shrinkage by taking exponent of the $P$ taken above and repeating the experiment 

## spcov with P = exp(1/ (common_samples))

```{r}
P <- exp(1/common_samples)
P[P==Inf] = 5
diag(P) <- 0
P[1:5,1:5]
```


```{r}
lam <- 0.1
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

```{r}
lam <- 0.3
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```


```{r}
lam <- 0.5
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

```{r}
lam <- 1
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```


## P  = 1/abs(S_{ij}) : Zou (2006)

```{r}
P <- 1/abs(cor_mat_pd)
P[P==Inf] = 1
diag(P) <- 0
P[1:5,1:5]
```

```{r}
lam <- 0.1
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

```{r}
lam <- 0.3
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```


```{r}
lam <- 0.5
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

## P with equal off-diagonal shrinkage

Now we consider the equal off-diagonal shrinkage as proposed as default in
Bien and Tibshirani.

```{r}
P <- matrix(1, 53, 53)
P[P==Inf] = 1
diag(P) <- 0
P[1:5,1:5]
```

```{r}
lam <- 0.1
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```


```{r}
lam <- 0.5
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```


```{r}
lam <- 1
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

## P being scale for normal distribution sd of r

```{r}
P <- sqrt((1 - cor_mat_pd^2)^2/common_samples)
P[P==Inf] = 1.5
diag(P) <- 0
P[1:5,1:5]
```

```{r}
lam <- 0.1
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```


```{r}
lam <- 0.5
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```


```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```


```{r}
lam <- 1
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

```{r}
lam <- 5
step.size <- 100
tol <- 1e-4
mm <- spcov(Sigma=cor_mat_pd+0.1*diag(1,53), S=cor_mat_pd+0.1*diag(1,53), lambda=lam * P,step.size=step.size, n.inner.steps=200,
thr.inner=0, tol.outer=tol, trace=1)
```

```{r}
spcor_mat <- cov2cor(mm$Sigma)
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(spcor_mat)[order_index, order_index],
      col=col, main=paste0("spcov:", expression(lambda), "=", lam), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```
