---
title: "Cross validation thresholding vs CorShrink"
author: "Kushal K Dey"
date: "5/25/2017"
output: html_document
---

In this script, we compare CorShrink with positive definite sparse covariance matrix estimation due to Rothman et al. We use the **PDSCE** package for this purpose.

## Data preparation

```{r}
library(CVTuningCov)
```

```{r}
gene <- "ENSG00000125414" ## update and rerun to see the plots for your gene
```

The number of samples per tissue 

```{r}
common_samples <- get(load("../output/common_samples.rda"))
```

```{r}
gene_names <- as.character(read.table(file = "../data/GTEX_V6/gene_names_GTEX_V6.txt")[,1])
gene_names_1 <- as.character(sapply(gene_names, function(x) return(strsplit(x, "[.]")[[1]][1])))
 
person_label=read.table("../data/GTEX_V6/person_identifier_labels_with_numbers.txt");
samples_id <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,1]

samples_person <- sapply(samples_id, function(x) return(paste0(strsplit(as.character(x), "-")[[1]][1:2], collapse ="-")))
tissue_labels <- read.table(file = "../data/GTEX_V6/samples_id.txt")[,3]

unique_persons <- unique(samples_person)
unique_tissues <- unique(tissue_labels)
```

```{r echo=FALSE, eval=TRUE}
order_index <- c();
U <- unique_tissues
order_index <- c(order_index, grep("Brain", U))
order_index <- c(order_index, grep("fibroblast", U))
order_index <- c(order_index, grep("EBV", U))
order_index <- c(order_index, grep("Spleen", U))
order_index <- c(order_index, grep("Whole Blood", U))
order_index <- c(order_index, grep("Muscle - Skeletal", U))
order_index <- c(order_index, grep("Liver", U))
order_index <- c(order_index, grep("Pancreas", U))
order_index <- c(order_index, grep("Stomach", U))
order_index <- c(order_index, grep("Kidney - Cortex", U))
order_index <- c(order_index, grep("Adrenal Gland", U))
order_index <- c(order_index, grep("Colon - Transverse", U))
order_index <- c(order_index, grep("Small Intestine - Terminal Ileum", U))
order_index <- c(order_index, grep("Heart - Atrial Appendage", U))
order_index <- c(order_index, grep("Heart - Left Ventricle", U))
order_index <- c(order_index, grep("Minor Salivary Gland", U))
order_index <- c(order_index, grep("Skin - Sun Exposed", U))
order_index <- c(order_index, grep("Skin - Not Sun Exposed", U))
order_index <- c(order_index, grep("Lung", U))
order_index <- c(order_index, grep("Ovary", U))
order_index <- c(order_index, grep("Thyroid", U))
order_index <- c(order_index, grep("Pituitary", U))
order_index <- c(order_index, grep("Testis", U))
order_index <- c(order_index, grep("Nerve - Tibial", U))
order_index <- c(order_index, grep("Breast - Mammary Tissue", U))
order_index <- c(order_index, grep("Adipose - Visceral", U))
order_index <- c(order_index, grep("Adipose - Subcutaneous", U))
order_index <- c(order_index, grep("Artery - Coronary", U))
order_index <- c(order_index, grep("Artery - Tibial", U))
order_index <- c(order_index, grep("Artery - Aorta", U))
order_index <- c(order_index, grep("Esophagus - Mucosa", U))
order_index <- c(order_index, grep("Vagina", U))
order_index <- c(order_index, grep("Cervix - Endocervix", U))
order_index <- c(order_index, grep("Esophagus - Gastroesophageal Junction", U))
order_index <- c(order_index, grep("Colon - Sigmoid", U))
order_index <- c(order_index, grep("Esophagus - Muscularis", U))
order_index <- c(order_index, grep("Cervix - Ectocervix", U))
order_index <- c(order_index, grep("Fallopian", U))
order_index <- c(order_index, grep("Prostate", U))
order_index <- c(order_index, grep("Uterus", U))
order_index <- c(order_index, grep("Bladder", U))
```

We load the sample correlation matrix and the CorShrink matrix.

## Sample correlation

```{r fig.width=10, fig.height=12}
cor_result <- get(load("../output/cor_tissues_non_ash_voom_pearson.rda"))
cor_result[is.na(cor_result)] = 0
ash_cor_result <- get(load("../output/ash_cor_only_voom_pearson_normal_gtex_tissues.rda"))
```

```{r}
numg <- match(gene, gene_names_1)
cor_mat <- diag(1, 53) + cor_result[,,numg]
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(cor_mat[order_index, order_index],
      col=col, main=paste0("sample corr: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
```

## CorShrink matrix

```{r}
numg <- match(gene, gene_names_1)
ash_cor_mat <- ash_cor_result[,,numg]
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(ash_cor_mat)[order_index, order_index],
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)

```

```{r}
cor_mat_pd <- Matrix::nearPD(cor_mat, corr=TRUE)
cor_mat_pd <- as.matrix(cor_mat_pd$mat)
```

## Hard thresholding 

```{r}
sigma <- hard.thresholding(Sigma = cor_mat_pd, c = 0.2)
```

```{r}
numg <- match(gene, gene_names_1)
ash_cor_mat <- ash_cor_result[,,numg]
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(sigma)[order_index, order_index],
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)

```

```{r}
sigma <- hard.thresholding(Sigma = cor_mat_pd, c = 0.3)
```

```{r}
numg <- match(gene, gene_names_1)
ash_cor_mat <- ash_cor_result[,,numg]
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(sigma)[order_index, order_index],
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)

```

```{r}
sigma <- hard.thresholding(Sigma = cor_mat_pd, c = 0.5)
```

```{r}
numg <- match(gene, gene_names_1)
ash_cor_mat <- ash_cor_result[,,numg]
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(sigma)[order_index, order_index],
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)

```

## Soft thresholding 

```{r}
sigma <- soft.thresholding(Sigma = cor_mat_pd, c = 0.2)
```

```{r}
numg <- match(gene, gene_names_1)
ash_cor_mat <- ash_cor_result[,,numg]
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(sigma)[order_index, order_index],
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)

```

```{r}
sigma <- soft.thresholding(Sigma = cor_mat_pd, c = 0.3)
```

```{r}
numg <- match(gene, gene_names_1)
ash_cor_mat <- ash_cor_result[,,numg]
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(sigma)[order_index, order_index],
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)

```


```{r}
sigma <- soft.thresholding(Sigma = cor_mat_pd, c = 0.5)
```

```{r}
numg <- match(gene, gene_names_1)
ash_cor_mat <- ash_cor_result[,,numg]
col=c(rev(rgb(seq(1,0,length=1000),1,seq(1,0,length=1000))),
      rgb(1,seq(1,0,length=1000),seq(1,0,length=1000)))
image(as.matrix(sigma)[order_index, order_index],
      col=col, main=paste0("corshrink: "), cex.main=2,
      xaxt = "n", yaxt = "n", zlim=c(-1,1))
axis(1, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)
axis(2, at = seq(0, 1, length.out = 53), labels = unique_tissues[order_index], las=2, cex.axis = 0.5)

```

## Softimpute + random CV

```{r}
library(softImpute)
impute_method <- "svd" ## other option tested is "als"
```

```{r}
person_tissue_genes <- get(load("../output/person_tissue_genes_voom.rda"))
```

```{r}
numg <- grep(gene, gene_names_1)
mat <- person_tissue_genes[,,numg]
matc=biScale(mat,col.scale=FALSE,row.scale=FALSE,trace=TRUE)
fits3=softImpute(matc,rank.max=25,lambda=1,type=impute_method)
fitted_mat <- complete(mat,fits3,unscale=TRUE)
```

```{r}
dim(fitted_mat)
```

```{r}
cv_val <- regular.CV(fitted_mat, method = "SoftThresholding")
```

This did not quite work as it would return `k.grid` input parameter. But 
most likely this is because none of these thresholding estimators are good fit.

## Conclusions

Both hard and soft thresholding estimators look for the same threshold for
all correlation values and is not flexible in taking into consideration the
standard error difference in correlations due to different number of samples.
So, they are not ideal comparisons to CorShrink.

